---
title: Chapter 5 Analysis Documentation
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    latex_engine: xelatex
    number_sections: yes
geometry: margin=0.75in
---

```{r project-setup}
setwd('C:/Users/meglarse/GitHub/larsen-dissertation/analyses/chpt4_LPS/')

# Load packages
require(vegan)
require(ggplot2)
require(reshape2)
require(phylobase)
require(ade4)
require(ape)
require(RColorBrewer)
require(adephylo)
require(pander)

# Load source code
source("./bin/CareStream-data-converter.R")
```

```{r global-opts}
```

# Project Summary

**Overview**: For this experiment, *Synechococcus* sp WH7803 was reanimated from cyropreservation and transferred to a gradient of N:P ratios in a regression design with 5 replicates. After standardizing the concentration of each culture, membranes were harvested, run out on gels, and stained using ProQ 300 Emerald stain. Each banding profile was converted to xy coordinates to identify the intensity of each band and relativize each sample to the *E. coli* O55:B5 LPS standard (Sigma).


**Collaborators**

J.T. Lennon, *Indiana University, Bloomington, IN*

**Citation**

**Project questions**

1. Does NL or PL alter LPS profiles on *Synechococcus* cell walls?
  - Figure 1: LPS gel of strains with image of ancestor in NL and PL conditions

2. How does resistance alter LPS profiles?
3. Does NL or PL change the composition of bands on the cell membrane?
4. How does resistance related to the infection profiles?
5. What are the consequences for interactions with phage?
6. Does phenotype result in cost?


\newpage
\tableofcontents
\newpage


# Data summary

## Summary of major results

\newpage

```{r convert-data}
## Read in all gel information
gel.metadata = read.csv("./data/gel_metadata.csv", header = TRUE)
unique(gel.metadata$gel.date)
head(gel.metadata)
str(gel.metadata)
gel.metadata$gel.id = as.character(gel.metadata$gel.id)
gel.metadata$lps.sample = as.character(gel.metadata$lps.sample)
gel.metadata$lane = as.character(gel.metadata$lane)
gel.metadata$batch = as.factor(gel.metadata$batch)

## Read in lps sample list
lps.list = read.csv("./data/lps_extracts.csv", header = TRUE)
head(lps.list)
str(lps.list)
lps.list$strain = as.character(lps.list$strain)
lps.list$strain.id = as.character(lps.list$strain.id)
lps.list$lps.sample = as.character(lps.list$lps.sample)

## Populate gel list with strain information
gel.dat = merge(gel.metadata,lps.list, by = c("batch","lps.sample"))
dim(gel.metadata);dim(lps.list);dim(gel.dat)

write.csv(gel.dat, "./data/gelmetadata-complete.csv",row.names = FALSE)

## Read in and convert profile data
g1 = carestream.convert("2016-5-22_1_20150525-gel1_Data")
g2 = carestream.convert("2016-5-22_1_20150525-gel2_Data")
g3 = carestream.convert("2016-5-22_1_20150525-gel3_Data")
g4 = carestream.convert("2016-5-22_1_20160413-gel1_Data")
g5 = carestream.convert("2016-5-22_1_20160413-gel2_Data")
g6 = carestream.convert("2016-5-22_1_20160413-gel3_Data")
g7 = carestream.convert("2016-5-22_1_20160413-gel4_Data")
g8 = carestream.convert("2016-5-22_1_20160420-gel1_Data")
g9 = carestream.convert("2016-5-22_1_20160420-gel2_Data")
g10 = carestream.convert("2016-5-22_1_20160420-gel3_Data")
g11 = carestream.convert("2016-5-22_1_20160425-gel1_Data")
g12 = carestream.convert("2016-5-22_1_20150114-gel1_Data")

profile = rbind(g1,g2,g3,g4,g5,g6,g7,g8,g9,g10,g11,g12)


# Convert data to appropriate structure
str(profile)

profile = as.data.frame(profile,stringsAsFactors = FALSE)
profile$MW = as.numeric(profile$MW)
profile$intensity = as.numeric(profile$intensity)
profile$rel.int = as.numeric(profile$rel.int)
profile$Band = as.numeric(profile$Band)

#write.csv(profile, "./data/gelbands.csv",row.names = FALSE)

## Keep all data
#lps.dat = merge(gel.dat,profile, by = c("gel.id","lane"), all = TRUE)

## Keep only strain data
lps.dat = merge(gel.dat,profile, by = c("gel.id","lane"))
MW = lps.dat[lps.dat$sample.type.x == "MW",]
EcoliR= lps.dat[lps.dat$lps.sample == "EcoliR",]
EcoliSm= lps.dat[lps.dat$lps.sample == "EcoliSm",]

lps.dat = lps.dat[lps.dat$sample.type.x != "MW" & lps.dat$sample.type.x != "std",]

```


### List of selected strains for analysis
```{r trim-data}
## Subset selected strains from the total data set
sel.strains = read.csv("./data/lps_selected.csv", header = TRUE)
pander(sel.strains[,c(1,5,6,7,8,9)])

sel.strains = sel.strains[,c(1,4,5)]
lps.dat2 = lps.dat[,c(1,10,11,12,13,14,16,17,19)]

lps.sel.dat.all = merge(sel.strains,lps.dat2, by = c("gel.id","strain.id"), all = TRUE)
lps.sel.dat = merge(sel.strains,lps.dat2, by = c("gel.id","strain.id"))

setdiff(lps.sel.dat.all$strain.id,lps.sel.dat$strain.id)

lps.sel.dat$lim <- as.character(lps.sel.dat$lim)
#unique((lps.sel.dat$strain.id))
```

```{r,results = 'hide', fig.height=3.75, fig.width=6, echo=FALSE, fig.cap="Treatment Chemostats"}
avg <- read.csv("./data/cstat-means.csv", header = T)

pdf(file = "./figures/chpt4_fig-TimeSeries.pdf", pointsize = 8, width = 5, height = 4)

day.start <- -125
day.end <- 170

#par(mfrow = c(1,2))
par(mar = c(5,6,3,1)+0.1)

plot(avg$day,avg$NmeanSI,
       xlim = c(day.start,day.end),ylim = c(1*10^4,10^9), log = "y", xaxt = 'n',yaxt = 'n',
       xlab = "Time (d)", ylab = expression(paste("Abundance (mL"^"-1",")")), 
     cex.lab = 1.25, font.lab = 1,        
     lty = 1, lwd = 2, type = "l", col = "black", bg = "white", 
       pch = 21, cex = 1)
  
  axis(1, cex.axis = 1.25,c(-120,-80,-40,0,40,80,120,160))
  ticks <- seq(4, 9, by=1)
  labels <- sapply(ticks, function(i) as.expression(bquote(10^ .(i))))
  axis(2, cex.axis = 1.25, at = c(10^4,10^5, 10^6, 10^7, 10^8, 10^9), labels = labels, las = 1)
  #abline(v = 0, col = "grey", lty = 2, lwd = 2)
  
## Add error bars to densities  
#  arrows(avg$day,avg$NmeanSI-avg$NseSI,avg$day,avg$NmeanSI+avg$NseSI,code=0,lwd=2)
#  arrows(avg$day,avg$PmeanSI-avg$PseSI,avg$day,avg$PmeanSI+avg$PseSI,code=0,lwd=2)

## Add polygon and points to graph
  
  if(day.start < 0){
    abline(v = 0, lwd = 2, lty = 1, col = "grey90")
    polygon(c(0,0,180,180),c(10^3,10^10,10^10,10^3),
     col = "grey90", border = NA)
  }
  
  points(avg$day,avg$NmeanSI,
         pch = 21, col = "steelblue",bg = "white",
         type = 'l',lwd = 3,cex = 1, lty = 1)
  points(avg$day,avg$PmeanSI,
         pch = 21, col = "forestgreen",bg = "white",
         type = 'l',lwd = 3,cex = 1, lty = 1)

## Phage densities  
#  points(avg$day,avg$NmeanP,
#         pch = 21, col = "black",bg = "black",
#         type = 'l',lwd = 1,cex = 1, lty = 1)

# Add isolate time points

iso.x = c(-125,-6,148,166)
isoN.y = c(104271961.3 ,50668258.2,22354875.8,73452512.2 )
isoP.y = c(84863414.33,38505264.08,23868855.67,68507860.57)

#abline(v = iso.x, lwd = 1.5, lty = 1, col = "grey25") 
points(iso.x, isoP.y, pch = 21, cex = 2.5, col = "black", bg = "black")
points(iso.x, isoN.y, pch = 21, cex = 2.5, col = "black", bg = "grey75")

#legend(c("Nlimited","P-limited","Strains"),col = c("steelblue","forestgreen"))

box(lwd=1.5)

dev.off()

```

```{r plot}
lps.sel.dat$strain.id <- as.character(lps.sel.dat$strain.id)
#lps.sel.dat <- lps.sel.dat[lps.sel.dat$MW > 14 & lps.sel.dat$MW < 82,]

lps.sel.dat$lim <- as.factor(lps.sel.dat$lim)
lps.sel.dat$tm.pt <- as.character(lps.sel.dat$tm.pt)

p1 = ggplot(data = lps.sel.dat[lps.sel.dat$lim == "N" & lps.sel.dat$tm.pt == "1",], 
            aes(strain.id, MW)) +
  #scale_color_gradient2(low = "blue", high = "black", guide = FALSE) +
  geom_point(pch = "-", size = 7) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5), 
        panel.border= element_rect(linetype = 1, fill = NA),
        axis.title = element_blank()) + ylim(10,65)

p2 = ggplot(data = lps.sel.dat[lps.sel.dat$lim == "N" & lps.sel.dat$tm.pt == "19",], 
            aes(strain.id, MW)) +
  geom_point(pch = "-", size = 7) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5), 
        panel.border= element_rect(linetype = 1, fill = NA),
        axis.title = element_blank()) + ylim(10,65)

p3 = ggplot(data = lps.sel.dat[lps.sel.dat$lim == "N" & lps.sel.dat$tm.pt == "21",], 
            aes(strain.id, MW)) +
  geom_point(pch = "-", size = 7) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5), 
        panel.border= element_rect(linetype = 1, fill = NA),
        axis.title = element_blank()) + ylim(10,65)

p4 = ggplot(data = lps.sel.dat[lps.sel.dat$lim == "P" & lps.sel.dat$tm.pt == "1",], 
            aes(strain.id, MW)) +
  geom_point(pch = "-", size = 7) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5), 
        panel.border= element_rect(linetype = 1, fill = NA),
        axis.title = element_blank()) + ylim(10,65)

p5 = ggplot(data = lps.sel.dat[lps.sel.dat$lim == "P" & lps.sel.dat$tm.pt == "19",], 
            aes(strain.id, MW)) +
  geom_point(pch = "-", size = 7) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5), 
        panel.border= element_rect(linetype = 1, fill = NA),
        axis.title = element_blank()) + ylim(10,65)

p6 = ggplot(data = lps.sel.dat[lps.sel.dat$lim == "P" & lps.sel.dat$tm.pt == "21",], 
            aes(strain.id, MW)) +
  geom_point(pch = "-", size = 7) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5), 
        panel.border= element_rect(linetype = 1, fill = NA),
        axis.title = element_blank()) + ylim(10,65)
  
# Read in reference strain information
ref <- as.matrix(read.delim("./data/2016-5-22_1_20150129_gel1_stdEval_Data.txt", sep = "\t", header = T))
MW <- cbind(rep("MW",5),ref[2:6,2],rep("MW",5))
Sm <- cbind(rep("Sm",21),ref[2:nrow(ref),5],rep("Standard",21))
R <- cbind(rep("R",5),ref[2:6,8],rep("Standard",5))
A <- cbind(rep("A",length(lps.sel.dat$MW[lps.sel.dat$strain.id == "101-7803-A"])),
            lps.sel.dat$MW[lps.sel.dat$strain.id == "101-7803-A"],
           rep("syn",length(lps.sel.dat$MW[lps.sel.dat$strain.id == "101-7803-A"])))
AR <- cbind(rep("AR",length(lps.sel.dat$MW[lps.sel.dat$strain.id == "232-7803R8"])),
            lps.sel.dat$MW[lps.sel.dat$strain.id == "232-7803R8"],
            rep("syn",length(lps.sel.dat$MW[lps.sel.dat$strain.id == "232-7803R8"])))
AR2 <- cbind(rep("AR2",length(lps.sel.dat$MW[lps.sel.dat$strain.id == "233-7803R8R21"])),
            lps.sel.dat$MW[lps.sel.dat$strain.id == "233-7803R8R21"],
            rep("syn",length(lps.sel.dat$MW[lps.sel.dat$strain.id == "233-7803R8R21"])))
ref <- rbind(MW,Sm,R,A,AR,AR2)

ref = as.data.frame(ref)
colnames(ref) <- c("Reference","MW","Type")
ref$MW <- as.numeric(as.character(ref$MW))
ref$Reference <- as.character(ref$Reference)
ref$Reference[ref$Reference == "A"] <- "101-7803-A"
ref$Reference[ref$Reference == "AR"] <- "232-7803R8"
ref$Reference[ref$Reference == "AR2"] <- "233-SR8R21"

p7 = ggplot(data = ref, aes(Reference, MW, col = Type)) +
  geom_point(pch = "-", size = 7) +
  scale_color_brewer(palette = "Dark2", guide = FALSE) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5), 
        panel.border= element_rect(linetype = 1, fill = NA),
        axis.title.x = element_blank()) + 
  ylab("Molecular Weight (kD)") + ylim(10,65)

p7$data$Reference <- factor(p7$data$Reference, levels = c("MW","Sm","R","101-7803-A","232-7803R8","233-SR8R21"))

#print(p7)

require(gridExtra)
pN = grid.arrange(p7,p1,p2,p3, nrow = 1, ncol = 4)
pP = grid.arrange(p7,p4,p5,p6, nrow = 1, ncol = 4)

#ggsave("./figures/chpt4_fig-bandprofilesN.pdf", plot = pN, dpi = 600, units = "in", width = 6, height = 2.5, scale = 2)
#ggsave("./figures/chpt4_fig-bandprofilesP.pdf", plot = pP, dpi = 600, units = "in", width = 6, height = 2.5, scale = 2)

# create matrix
#lps.sel.dat$MW1 = as.factor(round(lps.sel.dat$MW, digits = 1))
#lps.mat = acast(lps.sel.dat, strain.id ~ MW1, value.var = "rel.int")



```


```{r}

## Plot reference and control strains together; then treatment strains


p1 = ggplot(data = lps.sel.dat[lps.sel.dat$trt == "C" & lps.sel.dat$lim == "N",], 
            aes(strain.id, MW)) +
  scale_color_gradient2(low = "grey80", high = "black", guide = FALSE) +
  geom_point(pch = "-", size = 7) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5), 
        panel.border= element_rect(linetype = 1, fill = NA),
        axis.title = element_blank()) + ylim(10,65)

p2 = ggplot(data = lps.sel.dat[lps.sel.dat$trt == "T" & lps.sel.dat$lim == "N",], 
            aes(strain.id, MW)) +
  scale_color_gradient2(low = "grey80", high = "black", guide = FALSE) +
  geom_point(pch = "-", size = 7) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5), 
        panel.border= element_rect(linetype = 1, fill = NA),
        axis.title = element_blank()) + ylim(10,65)

p3 = ggplot(data = lps.sel.dat[lps.sel.dat$trt == "C" & lps.sel.dat$lim == "P",], 
            aes(strain.id, MW)) +
  scale_color_gradient2(low = "grey80", high = "black", guide = FALSE) +
  geom_point(pch = "-", size = 7) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5), 
        panel.border= element_rect(linetype = 1, fill = NA),
        axis.title = element_blank()) + ylim(10,65)

p4 = ggplot(data = lps.sel.dat[lps.sel.dat$trt == "T" & lps.sel.dat$lim == "P",], 
            aes(strain.id, MW)) +
  scale_color_gradient2(low = "grey80", high = "black", guide = FALSE) +
  geom_point(pch = "-", size = 7) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5), 
        panel.border= element_rect(linetype = 1, fill = NA),
        axis.title = element_blank()) + ylim(10,65)

grid.arrange(p1,p2,p3,p4, ncol = 2, nrow = 2)

```

\newpage

```{r}
bin.list2 <- c(); no.bands.list <- c(); strain.list <- c(); mean.rel.list <- c()

low = 18.25 - 0.68
high = 18.25 + 0.68

for(i in unique(lps.sel.dat$strain)){
  tmp <- lps.sel.dat[lps.sel.dat$strain == i,]
  
  
  start = tmp$MW[which(tmp$Band == max(tmp$Band))]
  
  for(j in 1:30){
    
  }
  
  strain.list <- append(strain.list, rep(as.character(i),(length(bin.list)-1)))
  
}
```

```{r modality 1}
ggplot(data = lps.sel.dat[lps.sel.dat$trt != "A",], aes(MW,rel.int, col = lim)) +
  stat_smooth(level = 0.99)+
  geom_point(pch = 20, size = 3) + xlim(20,60)+ ylim(0,0.4)+
  facet_wrap(~ tm.pt, ncol = 1)+
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5), 
        panel.border= element_rect(linetype = 1, fill = NA),
        axis.title = element_blank())

```

## Strain summary
```{r}
# calculate average information
#lps.sel.dat = lps.sel.dat[lps.sel.dat$MW > 14,]

range.list = c(); band.list = c(); strain.list = c();lim.list = c(); tm.list<-c(); trt.list <- c(); min.list <- c(); max.list <- c(); cID.list <- c(); mean.list <- c(); mode.list <- c()

#lps.sel.dat <- lps.sel.dat[lps.sel.dat$strain.id != "231-8101-A", ]

for(i in unique(lps.sel.dat$strain.id)){
  tmp = lps.sel.dat[lps.sel.dat$strain.id == i,]
  no.bands = dim(tmp)[1]
  range = max(tmp$MW)-min(tmp$MW)
  min = min(tmp$MW)
  max = max(tmp$MW)
  
  range.list = append(range.list, range)
  mean.list = append(mean.list, mean(tmp$MW))
  mode.list = append(mode.list, mode(tmp$MW))
  band.list = append(band.list, no.bands)
  strain.list = append(strain.list,as.character(i))
  lim.list = append(lim.list, as.character(tmp[1,5]))
  tm.list = append(tm.list, tmp[1,7])
  trt.list = append(trt.list, as.character(tmp[1,4]))
  min.list <- append(min.list, min)
  max.list <- append(max.list, max)
  cID.list <- append(cID.list, tmp[1,6])
}


avg.inf = data.frame(trt.list, lim.list,cID.list, tm.list,strain.list, band.list, range.list, min.list, max.list, mean.list, mode.list)
#colnames(avg.inf) = c("Limitation","Strain","No. bands", "MW Range")

write.csv(avg.inf,file = "./data/avgstraininfo.csv", row.names = FALSE)

avg.inf2 <- avg.inf[avg.inf$strain.list != "232-7803R8" & avg.inf$strain.list != "233-7803R8R21" ,]
table1 = acast(avg.inf2, lim.list ~ tm.list, mean, value.var = "band.list")

p = ggplot(avg.inf2, aes(x=lim.list:tm.list, band.list, fill = lim.list, shape = tm.list)) +
  geom_boxplot(alpha = 0.4)+
  geom_point(size = 2, alpha = 0.7) +
  scale_fill_brewer(guide = FALSE, palette = "Greys") +
  theme_classic() +
  xlab("Limitation") + ylab("Number of Bands per strain") +
  theme(panel.border= element_rect(linetype = 1, fill = NA))
print(p)
#ggsave("./figures/chpt4_fig-NoBandsBoxplot.pdf", plot = p, width = 4, height = 4, units = "in", pointsize = 6)

p2 = ggplot(avg.inf2, aes(x=lim.list:tm.list, range.list, fill = lim.list, shape = tm.list)) +
  geom_boxplot(alpha = 0.4)+
  geom_point(size = 2, alpha = 0.7) +
  scale_fill_brewer(guide = FALSE, palette = "Greys") +
  theme_classic() +
  xlab("Limitation") + ylab("Molecular Weight range") +
  theme(panel.border= element_rect(linetype = 1, fill = NA))
print(p2)
#ggsave("./figures/chpt4_fig-BandRangeBoxplot.pdf", plot = p2, width = 4, height = 4, units = "in", pointsize = 6)
```

```{r anova}
avg.inf <- avg.inf[avg.inf$lim.list != "A",]
str(avg.inf)
avg.inf$tm.list <- as.numeric(as.character(avg.inf$tm.list))

fit1 <- aov(band.list ~ tm.list * lim.list, data = avg.inf)
summary(fit1)
TukeyHSD(fit1)

fit2 <- aov(range.list ~ tm.list * lim.list, data = avg.inf)
summary(fit2)
#TukeyHSD(fit2)


```

\newpage


# Profile clustering
In a previous study, we challenged host and phage strains in a time-shift experiment. Using the data pertaining to the selected strains in this study, we were able to compare how LPS profiles of *Synechococcus* strains are related to resistance to phage infection.

## Summary of major results
1. 

\newpage

```{r, eval = FALSE}
data(RFLPdata)
res <- RFLPdist(RFLPdata)
plot(hclust(res[[1]]), main = "Euclidean distance")
 
par(mfrow = c(1,2))
plot(hclust(RFLPdist(RFLPdata, nrBands = 3)), cex = 0.7)
RFLPplot(RFLPdata, nrBands = 3, mar.bottom = 6, cex.axis = 0.8)
 
data(RFLPref)
RFLPrefplot(RFLPdata, RFLPref, nrBands = 6, cex.axis = 0.8)
 
```

## Clustering Examples with RFLPtools

RFLPtools is an analysis package for RFLP data that calculates presence/absence of bands for groups of samples with the same number. 

```{r}
require(RFLPtools)
lps.mat.dat = lps.sel.dat[,c(2,8:10)]
colnames(lps.mat.dat) = c("Sample","Bands","MW","rel.int")
lps.mat.dat$Sample = as.character(lps.mat.dat$Sample)

# RFLPtools only allows comparison of strains with the same number of bands
par(mfrow = c(1,2))
plot(hclust(RFLPdist(lps.mat.dat, nrBands = 6)), cex = 0.7) 
RFLPplot(lps.mat.dat, nrBands = 6, mar.bottom = 6, cex.axis = 0.8)

par(mfrow = c(1,2))
plot(hclust(RFLPdist(lps.mat.dat, nrBands = 8)), cex = 0.7) 
RFLPplot(lps.mat.dat, nrBands = 8, mar.bottom = 6, cex.axis = 0.8)

par(mfrow = c(1,2))
plot(hclust(RFLPdist(lps.mat.dat, nrBands = 10)), cex = 0.7) 
RFLPplot(lps.mat.dat, nrBands = 10, mar.bottom = 6, cex.axis = 0.8)


```

## Clustering profiles

```{r binning1}
# Create band bins using 7803R8

## read in data 
dat = as.matrix(read.delim("./data/2016-5-22_1_20160413-gel2_Data.txt", header = T, sep = "\t"))
dat <- dat[,26:28]
colnames(dat) <- dat[1,]
dat <- dat[-c(1,20:21),]
class(dat) = "numeric"

bin.list <- c()

for(i in 1:c(nrow(dat))-1){
  bin <- mean(c(dat[i,1], dat[i+1,1]))
  bin.list <- append(bin.list, bin)
}

#bin.list

## bin.list to categorize the number of bands and mean relative intensity within a bin for each strain

bin.list2 <- c(); no.bands.list <- c(); strain.list <- c(); mean.rel.list <- c()

lps.mat.dat <- na.omit(lps.mat.dat)

for(i in unique(lps.mat.dat$Sample)){
  tmp <- lps.mat.dat[lps.mat.dat$Sample == i,]
  
  for(j in 1:(length(bin.list)-1)){
    no.bands <- length(tmp$MW[tmp$MW < bin.list[j] & tmp$MW > bin.list[j+1]])
    mean.rel <- mean(tmp$rel.int[tmp$MW < bin.list[j] & tmp$MW > bin.list[j+1]])
    no.bands.list <- append(no.bands.list, no.bands)
    bin.list2 <- append(bin.list2, j)
    mean.rel.list <- append(mean.rel.list, mean.rel)
  }
  
  strain.list <- append(strain.list, rep(as.character(i),(length(bin.list)-1)))
  
}

bins <- data.frame(strain.list, bin.list2, no.bands.list, mean.rel.list)

## Replace NaN with 0
bins$mean.rel.list[bins$mean.rel.list == "NaN"] <- 0

#write.csv(bins, file = "./data/bins.csv", row.names = FALSE)

## create matrix
lps.mat <- acast(bins, strain.list ~ bin.list2,value.var = "mean.rel.list")

## create rel abundance distance matrix
lps.dist <- vegdist(decostand(lps.mat, method = "log"), method = "bray", na.rm = TRUE)


## create PA matrix
dataPA <- (lps.mat > 0)*1 
samplePA.dist = vegdist(decostand(dataPA, method = "log"), method = "jaccard")

```

```{r modality2}

bins <- read.csv("./data/bins.csv", header = T)
bins <- bins[no.bands.list >0,]

ggplot(data = bins[bins$lim != "A",], aes(bin, rel.int, col = lim)) +
#  scale_color_brewer(palette = "Blues") +
  geom_point(pch = 20, size = 2)+
  facet_wrap(~ time, nrow = 3)+
  stat_summary(fun.y = mean, geom = "line", size= 3)+


```



\newpage



```{r clustering-figure}
dev.off()

# heirarchical clustering algorithm for distance matrix
hc <- hclust(lps.dist, method = "complete")
hcp <- as.phylo(hc)

#Plot parameters
par(mar=c(1,1,1,1) + 0.2)
mypalette <- colorRampPalette(brewer.pal(9, "YlOrRd"))
mypalette2 <- colorRampPalette(brewer.pal(9, "YlGnBu"))


#nj.tree <- nj(lps.dist)

# Define the Outgroup
#outgroup <- match("231-8101-A", nj.tree$tip.label)

# Create a Rooted Tree {ape}
#nj.rooted <- root(nj.tree, outgroup, resolve.root = TRUE)
#nj.rooted$edge.length <- nj.rooted$edge.length + min(nj.rooted$edge.length)
#nj.rooted <- drop.tip(nj.rooted, "231-8101-A")

# Plot the Rooted Tree{ape}
#pdf(file="./figures/chpt4_fig-NJprofiletree.pdf",width = 3, height = 6, pointsize = 8)
#par(mar = c(1,1,2,1) + 0.1)
#plot.phylo(nj.rooted, main = "Neigbor Joining Tree", "phylogram", use.edge.length = FALSE,
#direction = "right", cex = 0.6, label.offset = 1)
#add.scale.bar(cex = 0.7)
#dev.off()


#par(mfrow = c(3,1))
# Which is the most appropriate agglomeration method
#plot(hclust(lps.dist, method = "complete"), hang = 0.1)
#plot(hclust(lps.dist, method = "single"), hang = 0.1)
#plot(hclust(lps.dist, method = "average"), hang = 0.1)

#dev.off()

# Save plot
#pdf(file="./figures/chpt4_fig-ProfSimComplete.pdf",width = 4, height = 7, pointsize = 7)
plot(as.phylo(hclust(lps.dist, method = "average")))
#dev.off()
```
```{r anosim, eval = TRUE}
rows = which(avg.inf$trt == "A")
dat = avg.inf[-rows,]
dat$strain.list <- as.character(dat$strain.list)
dat$trt.list <- droplevels(dat$trt.list)
dat$lim.list <- droplevels(dat$lim.list)
#droplevels(dat$cID)

bins = bins[bins$strain.list != "233-7803R8R21",]
bins = bins[bins$strain.list != "101-7803-A",]
bins = bins[bins$strain.list != "232-7803R8",]
bins = bins[bins$strain.list != "231-8101-A",]

lps.mat <- acast(bins, strain.list ~ bin.list2,value.var = "mean.rel.list")
lps.dist <- vegdist(decostand(lps.mat, method = "log"), method = "bray", na.rm = TRUE)
#lps.dist <- vegdist(lps.mat, method = "bray", na.rm = TRUE)

lps.ano <- anosim(lps.dist, grouping = c(dat$lim.list), permutations = 999)
summary(lps.ano)
plot(lps.ano)
```

### Observations
1. **Single** linkage, also known as nearest neighbor, is defined as the distance between the closest pair of objects.
2. **Complete** linkage, also known as farthest neighber, calculates distance between the farthest pair.
3. **Average** linkage considers the average of distances between all pairs.

\newpage

## Resistance
```{r bor-dataFormatting}
# Read in data file
inf.mat <- as.matrix(read.csv("./data/bor.csv"))         

# Rearrange data file for the appropriate analyses
inf.mat <- t(inf.mat)
inf.strains <- inf.mat[8:nrow(inf.mat),6]
p.dat <- inf.mat[1:6,7:ncol(inf.mat)]
inf.mat <- inf.mat[-c(1:7),-c(1:6)]
rownames(inf.mat) <- inf.strains
class(inf.mat) <- "numeric"

# Subset the infection matrix for only the sequenced strains
index2 <- which(rownames(inf.mat)  %in% unique(lps.sel.dat$strain))
WH7803 <- inf.mat[rownames(inf.mat)=="WH7803",]
inf.mat2 <- inf.mat[index2,]
inf.mat2 <- rbind(inf.mat2,WH7803)
inf.mat2 <- inf.mat2[-c(27,31),-c(95,96)]
class(inf.mat2) = "numeric"
```

```{r resistance}
# Function to calculate resistance
infect.prop<-function(x){
  x <- x[!is.na(x)] # remove NA's
  length(x[x == 1])/length(x)
	}

inf.dat <- matrix(NA,nrow = length(inf.strains), ncol=4)

# Calculate resistance
#rownames(inf.dat) <- inf.strains
inf.dat[,1] <- inf.strains
inf.dat [,2] <- round(apply(inf.mat,1,infect.prop),digits = 2)
inf.dat[1:85,3] <- round(apply(inf.mat[1:85,2:46],1,infect.prop),digits = 2)
inf.dat[88:225,3] <- apply(inf.mat[88:nrow(inf.mat),47:96],1,infect.prop)
inf.dat[1:85,4] <- apply(inf.mat[1:85,47:96],1,infect.prop)
inf.dat[88:225,4] <- apply(inf.mat[88:nrow(inf.mat),2:46],1,infect.prop)
names(inf.dat) <- c(NULL,NULL)

#Subset calculated  matrix with the sequenced strain names
#WH7803 <- inf.dat[inf.dat[,1]=="WH7803"]
index <- which(inf.dat[,1]  %in% unique(lps.sel.dat$strain)) 
inf.dat2 <- inf.dat[index,]
#inf.dat2 = inf.dat2[-c(27,31),]   # remove 12P2S5, 21P2S4 due to lack of inf data
#inf.dat2 <- rbind(WH7803,inf.dat2)
rownames(inf.dat2) <- NULL
colnames(inf.dat2) = c("strain","Average","Home","Away")
#write.csv(inf.dat2,file = "./supporting-files/data/BOR.csv",row.names=FALSE)
rownames(inf.dat2) <- inf.dat2[,1]
inf.dat2 <- inf.dat2[,-1]
class(inf.dat2)="numeric"

```

### Observations


\newpage
```{r clustering-figure1, eval = FALSE}
# Need to change the row names to get them to match up

inf.dat3 = merge(sel.strains,inf.dat2, by.x = "strain", by.y = 0)
rownames(inf.dat2) = inf.dat2[,3]


# heirarchical clustering algorithm for distance matrix
hc <- hclust(lps.dist, method = "complete")
hcp <- as.phylo(hc)

#P lot parameters
par(mar=c(1,1,1,1) + 0.2)
mypalette <- colorRampPalette(brewer.pal(9, "YlOrRd"))
mypalette2 <- colorRampPalette(brewer.pal(9, "YlGnBu"))


# Save plot
#pdf(file="./supporting-files/figures/fig-pub_traits-res.pdf",width = 3, height = 6)

x <- phylo4d(x = hcp, inf.dat3)
table.phylo4d(x, treetype = "phylo", symbol = "colors", show.node = TRUE, 
              cex.label = 0.35, scale = FALSE, use.edge.length = TRUE,
              edge.color = "black", edge.width = 1, box = FALSE, grid = TRUE,
              col= mypalette2(25), 
              # col = gray.colors(15, start = 0.3, end = 0.9, gamma = 2.2, alpha = NULL),
              pch = 15, cex.symbol = 1.25,
              ratio.tree = 0.25, cex.legend = 1.25, center = FALSE) 

#dev.off()

```


```{r clustering-figure2, eval = FALSE}
# Resistance + mutations
pdf(file="./supporting-files/figures/fig-pub_traits2.pdf",width = 5, height = 6)

x <- phylo4d(x = hcp, tree.dat2)
table.phylo4d(x, treetype = "phylo", symbol = "colors", show.node = TRUE, 
              cex.label = 0.35, scale = FALSE, use.edge.length = TRUE,
              edge.color = "black", edge.width = 2, box = FALSE, grid = TRUE,
              col= mypalette2(50), 
              # col = gray.colors(15, start = 0.3, end = 0.9, gamma = 2.2, alpha = NULL),
              pch = 15, cex.symbol = 1,
              ratio.tree = 0.20, cex.legend = 1, center = FALSE) 

dev.off()
```

\newpage

### Observations

\newpage

```{r dist-stats, eval = FALSE}
require(ade4)

#check dimensions of matrices
dim(breseq);dim(inf.mat2)
setdiff(rownames(breseq),rownames(inf.mat2))
setdiff(rownames(inf.mat2),rownames(breseq))

rmv=c(22,28,31,33,34)
breseq = breseq[-rmv,]
#inf.mat2= inf.mat2[-32,]
#inf.dat2 = inf.dat2[-32,]

# Make a distance matrix for the mutational data
breseq.dist <- vegdist(breseq, method="gower", na.rm = TRUE)
inf.dist <- vegdist(inf.mat2, method="gower", na.rm = TRUE)

# perform Mantel Test between two distance matrices
man.test <- mantel.randtest(breseq.dist, inf.dist, nrepet = 999)

plot(man.test)
cols = c(2:8)
dat <- tree.dat[-32,cols]
dat[,4] = droplevels(dat[,4])
adonis(breseq.dist ~ dat$Average*dat$lim*dat$trt, permutations = 999)
adonis(inf.dist ~ dat$Average*dat$lim*dat$trt, permutations = 999)
```


# Profile Oridination
## Summary of major results

\newpage

## Unweighted ordination
```{r PA-ordination}

# Principal Coordinates Analysis
lps_pcoaPA <- cmdscale(samplePA.dist,k=3,eig=TRUE,add=FALSE) 

# Percent Variance Explained Using PCoA (Axis 1,2,3)
explainvar1 <- round(lps_pcoaPA$eig[1]/sum(lps_pcoaPA$eig)*100,1) 
explainvar2 <- round(lps_pcoaPA$eig[2]/sum(lps_pcoaPA$eig)*100,1)
explainvar3 <- round(lps_pcoaPA$eig[3]/sum(lps_pcoaPA$eig)*100,1)
  
pcoap <- merge(as.data.frame(lps_pcoaPA$points),lps.list,by.x=0, by.y = "strain.id")
rownames(pcoap) <- rownames(lps_pcoaPA$points)
#write.csv(file="../supporting_files/data/pcoap-UnweightedOrdination.csv",pcoap)

#pdf(file = "../supporting_files/figures/pub-figure_chpt4-PCoA-UnweightedOrdination.pdf", width = 5, height = 5, pointsize = 10)

# Plot Parameters
par(mar=c(5,5,1,1))#, oma=c(3,1,1,1)+0.1 )
x.dim <- c(min(pcoap$V1)-(max(pcoap$V1)*0.1),max(pcoap$V1)+(max(pcoap$V1)*0.1))
y.dim <- c(min(pcoap$V2)-(max(pcoap$V2)*0.1),max(pcoap$V2)+(max(pcoap$V2)*0.1))

# Initiate Plot
plot(pcoap$V1, pcoap$V2, xlab=paste("PCoA Axis 1 (",explainvar1, "%)", sep="")
    , ylab=paste("PCoA Axis 2 (",explainvar2, "%)", sep=""), 
    xlim=x.dim, ylim=y.dim, pch=16, cex=2.0, type="n",xaxt="n",
    yaxt="n", cex.lab=1.5, cex.axis=1.2)
axis(side=1, las=1)   
axis(side=2, las=1)    
abline(h=0, lty="dotted")  
abline(v=0, lty="dotted")
box(lwd=2)

# Make Plot Symbols in PCoA Reflect Treatment
 
#mol.shape <- rep(NA, dim(pcoap)[1])
#  for (i in 1:length(mol.shape)){
#    if (pcoap$lim[i] == "N"){mol.shape[i] = 21}
#    else {mol.shape[i] = 22}
#	}

#slope.color <- rep(NA, dim(pcoap)[1])
#  for (i in 1:length(slope.color)){
#    if (pcoap$lim[i] == "N") {slope.color[i] = "black"}
#    else {slope.color[i] = "white"}  
#	} 

# Add Points & Ellipses
points(pcoap$V1, pcoap$V2, pch = 20, cex = 2.5, 
       bg = "black", col = "gray25")
#points(pcoap$V1, pcoap$V2, pch = mol.shape, cex = (pcoap$time+1)*0.2, 
#       bg = slope.color, col = "gray25")
#text(pcoap$V1, pcoap$V2, labels = pcoap$time, pos = 3)  
text(pcoap$V1, pcoap$V2, labels = pcoap$strain, pos = 1) 

#dev.off()


```


\newpage

## *********************Weighted ordination
```{r CommPCoA, eval = TRUE}
# Create distance matrix with band relative abundance
#lps.dist <- vegdist(lps.mat, method = "bray", na.rm = TRUE)
lps.dist <- vegdist(decostand(cstat.mat, method = "log"), method = "bray", na.rm = TRUE)

# Principle Coordinates Analysis
lps_pcoa <- cmdscale(lps.dist,k=3,eig=TRUE,add=FALSE)


# Percent Variance Explained Using PCoA (Axis 1,2,3)
explainvar1 <- round(lps_pcoa$eig[1]/sum(lps_pcoa$eig)*100,1) 
explainvar2 <- round(lps_pcoa$eig[2]/sum(lps_pcoa$eig)*100,1)
explainvar3 <- round(lps_pcoa$eig[3]/sum(lps_pcoa$eig)*100,1)
  
pcoap <- merge(as.data.frame(lps_pcoa$points),lps.list,by.x=0, by.y = "strain.id")
rownames(pcoap) <- rownames(lps_pcoa$points)

# save figure input data
#write.csv(pcoap, "./supporting-files/data/chpt4_commPCoA.csv",row.names = FALSE)

# Save figure
#pdf(file = "./figures/chpt3_commPCoA.pdf", width = 5, height = 5, pointsize = 10)

# Plot Parameters
par(mar=c(5,5,2,1))#, oma=c(3,1,1,1)+0.1 )
x.dim <- c(min(pcoap$V1)-(max(pcoap$V1)*0.1),max(pcoap$V1)+(max(pcoap$V1)*0.1))
y.dim <- c(min(pcoap$V2)-(max(pcoap$V2)*0.1),max(pcoap$V2)+(max(pcoap$V2)*0.1))

# Initiate Plot
plot(pcoap$V1, pcoap$V2, xlab=paste("PCoA Axis 1 (",explainvar1, "%)", sep="")
    , ylab=paste("PCoA Axis 2 (",explainvar2, "%)", sep=""), 
    xlim=x.dim, ylim=y.dim, pch=16, cex=2.0, type="n",xaxt="n",
    yaxt="n", cex.lab=1.5, cex.axis=1.2, bty='L')
axis(side=1, las=1)   
axis(side=2, las=1)    
abline(h=0, lty="dotted",col = "grey50")  
abline(v=0, lty="dotted",col = "grey50")
box(lwd=2)


# Add Points & Ellipses
points(pcoap$V1[pcoap$lim == "N" & pcoap$trt == "C"], 
       pcoap$V2[pcoap$lim == "N" & pcoap$trt == "C"], 
       pch = 21, cex = 3.5, bg = "steelblue", col = "gray25")
points(pcoap$V1[pcoap$lim == "N" & pcoap$trt == "T"], 
       pcoap$V2[pcoap$lim == "N" & pcoap$trt == "T"], 
       pch = 22, cex = 3.5, bg = "steelblue", col = "gray25")
points(pcoap$V1[pcoap$lim == "P" & pcoap$trt == "C"], 
       pcoap$V2[pcoap$lim == "P" & pcoap$trt == "C"], 
       pch = 21, cex = 3.5, bg = "forestgreen", col = "gray25")
points(pcoap$V1[pcoap$lim == "P" & pcoap$trt == "T"], 
       pcoap$V2[pcoap$lim == "P" & pcoap$trt == "T"], 
       pch = 22, cex = 3.5, bg = "forestgreen", col = "gray25")

#text(pcoap$V1, pcoap$V2, labels = pcoap$strain, pos = 1) 

legend("bottomleft",pch = c(19,15), c("N-limited","P-limited"),bty = 'n', col = c("steelblue","forestgreen"),cex = 1.25)

#dev.off()
```

```{r PERMANOVA, eval = TRUE}
rows = which(pcoap$trt == "A")
dat = pcoap
dat$Row.names <- as.character(dat$Row.names)
#droplevels(dat$trt)

bins = bins[bins$strain.list != "233-7803R8R21",]
bins = bins[bins$strain.list != "101-7803-A",]
bins = bins[bins$strain.list != "232-7803R8",]
bins = bins[bins$strain.list != "231-8101-A",]

lps.mat <- acast(bins, strain.list ~ bin.list2,value.var = "mean.rel.list")
lps.dist <- vegdist(decostand(lps.mat, method = "log"), method = "bray", na.rm = TRUE)
#lps.dist <- vegdist(lps.mat, method = "bray", na.rm = TRUE)

fit <- adonis(lps.dist ~ lim*trt, strata = dat$cID, data = dat ,permutations = 999)


dataPA <- (lps.mat > 0)*1 
#samplePA.dist = vegdist(dataPA, method = "jaccard")
samplePA.dist = vegdist(decostand(dataPA, method = "log"), method = "jaccard")

adonis(samplePA.dist ~ lim*trt*tm.pt, strata = dat$cID, data = dat ,permutations = 999)

```

#### Observations



\newpage



# Additional information

All analyses in this document were completed using `r sessionInfo()$R.version$version.string`. 

Add list of packages and versions used for analyses
Add list of citations for packages and/or other code chunks